---
output: beamer_presentation
---

## Title slide

Thank you for coming. I'm grateful for being able to present in this session. 

## Overview of talk

- A little about SWOT satellite mission this work is supporting
- Bayesian model for river discharge using satellite observations
- Proxy dataset for testing this model
- Results on proxy dataset

## Data assimilation, Bayesian, large dataset

The real large dataset is not yet in operation, I'm going to show results from a proxy, old-fashioned dataset

## Introduction: SWOT satellite mission

- who: NASA, CNES
- what: global topography using wide-swath radar interferometry
    - recurrence interval (?)
- when: 2021 - 2026(?)

## Measurements

- height via radar interferrometry
- width via image classification?
- slope via height / distance
- dA via integration: $\int_{w' < w} w \cdot dh $

## Estimation approch: Mass conserved flow law inversion (McFLI)

- Flow law: typically Manning's equation
    - also AMHG (Gleason citation)
- Mass conservation at the reach or multi-reach scale

## The model I use

- Manning's equation

$$
Q = \frac{1}{n}A^{5/3}W^{-2/3}S^{1/2}
$$


- log transform (Bayesians *love* log space)

$$
\log Q  =  - \log n  + \frac{5}{3} \log A - \frac{2}{3} \log W + \frac{1}{2} \log S
$$

- Adjust for SWOT observables

$$
\log Q  =  - \log n  + \frac{5}{3} \log (A_0 + \delta A) - \frac{2}{3} \log W + \frac{1}{2} \log S
$$



- Put indices on it

$$
\log Q_{i t}  =  - \log n  + \frac{5}{3} \log (A_{0, i} + \delta A_{it}) - \frac{2}{3} \log W_{it} + \frac{1}{2} \log S_{it}
$$

- Mass conservation removes one of the Q indices

$$
\log Q_{t}  =  - \log n  + \frac{5}{3} \log (A_{0, i} + \delta A_{it}) - \frac{2}{3} \log W_{it} + \frac{1}{2} \log S_{it}
$$

- Rearrange it, acknowledge error

$$
\frac{2}{3} \log W_{it} - \frac{1}{2} \log S_{it}  =   - \log n  + \frac{5}{3} \log (A_{0, i} + \delta A_{it}) - \log Q_t + \epsilon_{it}
$$

- Now can write the likelihood

$$
f(y_{it} | n, A_{0,i}, \delta A_{it}, Q_t) \sim N(\mu = RHS, \sigma)
$$

- Of course, we're not quite finished, since $\delta A$ is measured. We want:

$$
f(y_{it}, \delta A_{it} | n, A_{0,i}, Q_t)
$$

- But that's just:

$$
f(y_{it}, \delta A_{it} | n, A_{0,i}, Q_t) = f(y_{it} | n, A_{0,i}, \delta A_{it}, Q_t) \cdot f(\delta A_{it} | n, A_{0,i}, Q_t)
$$

First part of rhs is boring, normal, familiar. But impossible to use without havning the second part. I'm going to focus on second part. 

## Part 2: Inferring $A_0$ from $\delta A$

First: Assume a distribution for $A$ at a particular cross-section.

$$
A \sim \log N(\mu_{\log A}, \sigma_{\log A})
$$

Second: for a clever adjustment of $\delta A$, $\mu_{\log A} = \log A_0$

Then:

$$
f_{\delta A | \sigma_A, A_0}(x |  \sigma_A, A_0) = \frac{1}{(x + A_0) \sqrt{2 \pi} \sigma_A} \exp{\big(-\frac{1}{2 \sigma_A^2}[\log (x + A_0) - \log A_0]^2\big)}, (x > -A_0)
$$

So simple! (This actually ignores autoregression component we use when aggregating data)

## Or: just do this:

- Find legacy datasets of $A$, $W$
- Split into $A_0$, $\delta A$
- Fit a line

## (Aside about HydroSWOT dataset)

- Collected by
- Map of stations
- Stats of observations
- Acknowledgment

## Results

First, a demonstration

```{r}
demoset1 <- test_full %>% 
  filter(xs == 6610000)
priors1 <- makeMlePriors(test_smry %>% filter(xs == 6610000))


plot(demoset1$dA, type = "l")
plot(demoset1$area_m2, type = "l")
```

```{r}
demoset1$dH <- with(demoset1, c(0, diff(dA)) / w_m)

ggplot(demoset1, aes(x = h_m, y = dA)) +
  geom_point()

ggplot(demoset1, aes(x = w_m / 2, y = cumsum(dH), 
                     xend = -w_m / 2, yend = cumsum(dH))) +
  geom_segment()

ggplot()
```

Demo 1: Animation of dA timeseries 

```{r}

ts_anim(demoset1$dA, demoset1$Date)
1
```




Demo 2: Add histogram to the right.

```{r}
hist_anim(demoset1$dA, demoset1$Date)
```


Demo 2: Add timesieries of A0 estimate

```{r}
# mlets1 <- mle_ts(logPost_ar1, data.frame(obs = demoset1$dA, demoset1$Date), p1 = c(6.4, 5))
# plot(mlets1[, 1])
# plot(demoset1$dA)

mlets2 <- mle_ts(logPost_ar1, data.frame(obs = demoset1$dA, demoset1$Date), p1 = c(6.4, 5), 
                 priors = priors1)
ts_anim(mlets2[, 1])
```


Demo 4: Add 2-d contour plot for A0, sigmalogA

```{r}
nllfun1 <- logPost_ar1(data.frame(obs = demoset1$dA, dates = demoset1$Date), 
                       priors = priors1)
grid1 <- nll_grid(xvec = seq(6.44, 7, length.out = 100L), 
                  yvec = 10^(seq(-2, -0.3, length.out = 100L))
                  
```


Demo 5: The same, now with prior distribution from HydroSWOT

## Overall results:

Prior distribution

tweenr to posterior distribution

Remove low-area-variability cases

## Why are low-area variability cases bad?

Static image of likelihood contours

## Recap and Conclusions

## Future work
